---
title: "Lab 11 - Smoking during pregnancy"
author: "Qilin Zhang"
date: "3/30/23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
library(psych)
library(skimr)
library(boot)
```

### Exercise 1

I see 13 variables in total. There are 6 numeric variables and 7 categorical variables. 
```{r check}
data(ncbirths)

describe(ncbirths)
skim(ncbirths)

# visualization

ncbirths %>%
  ggplot(aes(x=weight))+
  geom_density()

ncbirths %>%
  ggplot(aes(x=gained))+
  geom_density()

ncbirths %>%
  ggplot(aes(x=visits))+
  geom_density()

ncbirths %>%
  ggplot(aes(x=weeks))+
  geom_density()
```


### Exercise 2&3

It seems like there is only a few means lower than 7.250462 in the bootstrapping sample. After accounting for the proportion of the means lower than 7.25, the p value with 15000 rep of bootstraping is 0.0005. As a result, we can conclude that the mean of child's weight is significantly lower than those in 1995.

```{r white_baby}
ncbirths_white <- ncbirths %>%
  filter(whitemom == "white")

mean(ncbirths_white$weight)

#bootstrap
boot_bir_white <- ncbirths_white %>%
  # specify the variable of interest
  specify(response = weight) %>%
  # generate 15000 bootstrap samples
  generate(reps = 15000, type = "bootstrap") %>%
  # calculate the mean of each bootstrap sample 
  calculate(stat = "mean")

boot_bir_white %>%
  summarize(lower = quantile(stat, 0.025),
            mean = mean(stat),
            upper = quantile(stat, 0.975))

boot_bir_white$stat <- boot_bir_white$stat - (mean(ncbirths_white$weight) - 7.43)
mean(boot_bir_white$stat)

boot_bir_white %>%
  summarize(lower = quantile(stat, 0.025),
            mean = mean(stat),
            upper = quantile(stat, 0.975))

boot_bir_white %>%
  filter(stat <= 7.250462) %>%
  nrow()/nrow(boot_bir_white)# p value = 0.0005


#visual
boot_bir_white %>%
  ggplot(aes(x=stat)) + 
  geom_histogram(binwidth = 0.01)+
  geom_vline(xintercept = 7.324958)+
  geom_vline(xintercept = 7.53526)+
  geom_vline(xintercept = 7.250462, color = "red")
  labs(title = "distribution for bootstrap means")

```

###exercise 4

one sample t-test suggests that the mean of child's weight is significantly lower than those in 1995.
```{r hypo_test}
t.test(x= ncbirths_white$weight,mu = 7.43)
```

###exercise 5 -6 

From the plots, it seems like the baby weight of non-smokers are higher than smokers. 
```{r boxplot}
ncbirths %>%
  ggplot(aes(x=habit, y = weight, color = habit)) + 
  geom_boxplot()

ncbirths_habitgiven <- ncbirths %>%
  filter(habit != "NA")

```

###exercise 7+8

Null hypothesis: There is no difference between baby weight with smoking and non-smoking parents among the population.
Alternative hypothesis: There is difference between baby weight with smoking and non-smoking parents among the population.

```{r}
ncbirths_habitgiven %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight),
            median_weight = median(weight))

```

###exercise 9

There is difference between baby weight with smoking and non-smoking parents among the population. In particular, the baby weight with smoking parents are significantly lower than those of the non-smoking parents. The confident interval represents that we have 95% of confident that the results t value will land between 0.45-4.16. 

```{r t-test_boot}

Smokers <- ncbirths_habitgiven %>%
  filter(habit=="smoker") %>%
  select(weight)
Smokers$id <- c(1:nrow(Smokers))

nonsmokers <- ncbirths_habitgiven %>%
  filter(habit=="nonsmoker") %>%
  select(weight)
nonsmokers$id <- c(1:nrow(nonsmokers))
smoking_boot <- merge(nonsmokers,Smokers,
                      by = "id",
                      all = T) %>%
  select(!id)

tstat <- function(data, i) {
  x <- data[i,1]
  y <- data[i,2]
  t.test(x, y)$statistic
}

set.seed(1234)
boot_result <- boot(smoking_boot, tstat, R=15000)
mean(boot_result$t)

#confident interval
quantile(boot_result$t,0.025)
quantile(boot_result$t,0.975)

# transfer t to p
t_to_p <- function(t, df) {
  p <- 2 * pt(abs(t), df = df, lower.tail = FALSE)
  return(p)
}
t_to_p(mean(boot_result$t) ,999-2)

t.test(smoking_boot$weight.x,smoking_boot$weight.y)

```



###exercise 11

It looks like cutoff is at 34 to 35.

```{r cutoff}
cutoff <- ncbirths_habitgiven%>%
  filter(mature=="younger mom")%>%
  select(mage)%>%
  max()%>%
  as.numeric()
cutoff

ggplot(ncbirths_habitgiven, aes(x = mage, fill = mature))+
  geom_histogram(binwidth = 0.5)+
  geom_vline(xintercept = cutoff+0.5)
```

###exercise 12+13

Null hypothesis:  the proportion of low birth weight babies is not different among mature or non-mature mothers
Alternative hypothesis:  the proportion of low birth weight babies is higher for mature mothers

It seems like there is no significant difference on the proportion of low birthweight babies between mature and non-mature mothers. Therefore, the original hypothesis in the question does not stand. 

```{r chisquare}

mature_lowweight <- ncbirths_habitgiven %>%
  select(c(mature,lowbirthweight)) %>%
  table()
#direct way
chisq.test(mature_lowweight,simulate.p.value = T,B = 10000)

chi_sq_stat <- function(data, indices) {
  tbl <- data[indices,]
  chisq <- chisq.test(tbl)$statistic
  return(chisq)
}

boot_results <- boot(data = mature_lowweight, statistic = chi_sq_stat, R = 10000)
# confident intervals
boot.ci(boot_results, type = "basic",conf = 0.95)

```
